### 2021-03-01 ch01_introduction/introduction.tex
Finished chapter 07
Finished chapter 01

- Why is climate change relevant for the society?

- Cite latest warming reports, also other variables (droughts, extreme events,
etc.)

- basic physics of greenhouse gases (vibration modes, etc.)

- definition of radiative forcing

- well-mixed greenhouse gases (water vapor, CO2, CH4) + sources

- long-lived vs. short-lived GHGs?

- aerosols + cooling effect

- natural vs. anthropogenic effects (D\&A) -> clear evidence that climate
change over 21st century is caused by human influence. Website:
(https://globalwarmingindex.org/, Haustein et al., 2017)

- remove description of emergent constraints from section 1.3, since this will
probably be included in this section.

- answer question: what is climate sensitivity? What is \ac{ECS}? What is
\ac{TCR}? What is \ac{CMIP}?

- One sentence on \ac{ML}.

- Mention IPCC assessments, older ones and upcoming AR6

- "climate model" is used as generic term which include \acp{AOGCM} and
\acp{ESM} (see \cref{subsec:02:numerical_climate_modeling}) (subcategories).
Since most modern climate models participating in CMIP are \acp{ESM} (or at
least have \ac{ESM} versions), the terms "climate model" and "ESM" are used
interchangeably in this thesis.

@@

This \namecref{sec:01:key_science_questions} will present the \textbf{AWESOME}
\namecrefs{enum:01:question_1}!!

@@

This thesis is structured as follows: \Cref{ch:02:scientific_background}
introduces the scientific background. This includes relevant literature that is
used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
over the contributions made to the \ac{ESMValTool}, an open-source software for
the analysis of \acp{ESM}. These contributions helped improving the routine
evaluation of \acp{ESM} which is useful for the entire scientific community and
lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:papers_ecs_tcr_assessment}
covers the assessment of climate sensitivity metrics like the \ac{ECS} or
\ac{TCR} in the latest generation of \acp{ESM} (\acs{CMIP}6). This work is
already published in two scientific publications \autocite{Bock2020,
  Meehl2020}. Since the \ac{ECS} and \ac{TCR} are considerably higher in this
new climate model generation, \cref{ch:05:paper_ecs} describes the assessment
of emergent constraints (a technique to reduce uncertainties in multi-model
climate projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS}
for these \acp{ESM}. The contents of this \namecref{ch:05:paper_ecs} are
published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
\Cref{ch:06:paper_gpp} focuses on a new method to reduce uncertainties in
multi-dimensional (gridded) multi-model projections of the future climate with
observations based on \ac{ML}. As an example, the method is applied to the
photosynthesis rate at the end of the \nth{21} century, which is already
published in the \emph{Journal of Geophysical Research: Biogeosciences}
\autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of
the results of this thesis and gives an outlook of possible future works.

>>>>

Climate change is one of the greatest challenges for humankind today with
far-reaching consequences for all human beings on planet Earth. The warming of
the climate system is \enquote{unequivocal} and \enquote{many of the observed
  changes are unprecedented over decades to millennia} \autocite{IPCC2014}. The
changing climate increases the \enquote{likelihood of severe, pervasive and
  irreversible impacts for people, species and ecosystems} with \enquote{mostly
  negative impacts for biodiversity, ecosystem services and economic
  development} and amplifies \enquote{risks for livelihoods and for food and
  human security} \autocite{IPCC2014}. Potential drivers for climate change
are all natural and anthropogenic substances and processes that may alter the
Earth's energy budget. The human influence on the climate system is clear: Ever
increasing emissions of \acp{GHG} since the end of the pre-industrial era
largely driven by economic and population growth led to atmospheric
concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O} that are
\enquote{unprecedented in at least the last 800000 years} \autocite{IPCC2014}.
The effects of these \ac{GHG} emissions and other anthropogenic drivers have
been \enquote{detected throughout the climate system and are extremely likely
  to have been the dominant cause of the observed warming since the
  mid-\nth{20} century} \autocite{IPCC2014}. The greenhouse effect is based on
the optical properties of the \acp{GHG}: While the atmosphere is mostly
transparent for the incoming solar (shortwave) radiation, the outgoing
infrared (longwave) radiation that is reflected from the planet's surface is
partly absorbed by the corresponding \ac{GHG} molecules through an excitation
of their corresponding vibrational modes and re-emitted isotropically. This
process traps energy near the surface and leads to a warming of the Earth's
surface and the lower atmosphere. The individual impact of the different
drivers of climate change measured with the so-called \emph{radiative
forcing}, which quantifies the change in energy fluxes caused by changes in
these drivers relative to pre-industrial conditions \autocite{IPCC2013}. The
total radiative forcing is positive (which corresponds to a warming) and its
largest contribution is caused by the increase in the atmospheric
concentration of \ac{CO2} due to fossil fuel emissions since the year 1750
\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
of the terrestrial biosphere due to photosynthesis. Since this land carbon
uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions in
today's climate \autocite{Friedlingstein2020}, this process substantially slows
down global warming. Other important anthropogenic drivers of climate change
are the emission of aerosols and land use/land cover changes. Apart from
anthropogenic drivers, there are also natural processes which impact the
climate system like changes in the solar activity or volcanic eruptions.
However, there is clear evidence that these natural drives alone cannot explain
the observed climate change \autocite{Haustein2017}.

To successfully mitigate the massive impacts of climate change, a first
important step is the understanding of climate change and its accurate
quantification. Extremely valuable tools in this context are climate models,
which allow us to simulate the behavior of the climate system under arbitrary
conditions without having to perform (ethnically questionable) experiments in
the real world. All around the world climate research institutes provide a
variety of different climate models. Many of them participate in the \ac{CMIP},
which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
\enquote{better understand past, present and future climate changes arising
  from natural, unforced variability or in response to changes in radiative
  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
provide crucial input for the international climate assessments given by the
\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
  as \acs{CMIP}6} will be used in the upcoming Sixth \acl{AR} (\acs{AR}6) of
the \ac{IPCC}, and their predecessor models from \acs{CMIP}5
\autocite{Taylor2012} have been successfully evaluated in the Fifth \acl{AR}
(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
allow the simulation of biological and chemical processes in addition to the
dynamics of the physical components of the Earth system, are also known as
\acp{ESM} and provide the most sophisticated simulations of the Earth's
climate. In this thesis, the terms \enquote{climate model} and
\enquote{\acl{ESM}} are used interchangeably since most modern models
participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.

Simulations that extrapolate state of the climate system into the future are
called \emph{projections} of the future climate. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with high ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. State-of-the-art techniques for this are so-called
\emph{emergent constraints}, which use a physically-based inter-model
relationship between an observable quantity of the Earth system and a target
variable to reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is climate model weighting based
on performance (\ie{} the distance of the climate models relative to
observational products) and interdependence (\ie{} the distance between the
climate models among each other) \autocite{Knutti2017a}. These techniques form
the baseline for the new analyses and results presented in this thesis.

@@

The aim of this thesis is to reduce uncertainties in multi-model climate
projections with observations by addressing the following three
\namecrefs{enum:01:question_1}:

@@

This thesis is structured as follows: \Cref{ch:02:scientific_background}
introduces the scientific background. This includes relevant literature that is
used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
over the contributions made to the \ac{ESMValTool}, an open-source software for
the analysis of \acp{ESM}. These contributions helped improving the routine
evaluation of \acp{ESM} which is useful for the entire scientific community and
lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:papers_ecs_tcr_assessment}
covers the assessment of climate sensitivity metrics like the \ac{ECS} or
\ac{TCR} in the latest generation of \acp{ESM} from \acs{CMIP}6. This work is
already published in two scientific publications \autocite{Bock2020,
  Meehl2020}. Since the \ac{ECS} and \ac{TCR} are considerably higher in this
new climate model generation, \cref{ch:05:paper_ecs} describes the assessment
of emergent constraints on the \ac{ECS} for these \acs{CMIP}6 models and
compares these to results derived from \acs{CMIP}5 models. The contents of this
\namecref{ch:05:paper_ecs} are published in \emph{Earth System Dynamics}
\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
reduce uncertainties in multi-dimensional (gridded) multi-model projections of
the future climate with observations based on \ac{ML}. As an example, the
method is applied to \ac{GPP} at the end of the \nth{21} century, which is
already published in the \emph{Journal of Geophysical Research: Biogeosciences}
\autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of
the results of this thesis and gives an outlook of possible future works.

### 2021-03-01 ch02_scientific_background/scientific_background.tex
Finished chapter 01

An important metric describing climate sensitivity is the equilibrium climate
sensitivity. It is defined as the change in \ac{GSAT} after an instantaneous
doubling of the atmospheric \ac{CO2} concentration from pre-industrial
conditions once the climate system reaches radiative equilibrium
\autocite{Bindoff2013}. Being already used in one of the first assessments of
the anthropogenic climate change, the \emph{Charney Report} from 1979
\autocite{Charney1979}, the equilibrium climate sensitivity is one of the
oldest metrics describing climate change. However, in practice this traditional
definition of is not useful. Running fully-coupled \acp{ESM} into equilibrium
is computationally expensive as it would require thousands of model years
\autocite{Rugenstein2020}.

@@

For this reason, the equilibrium climate sensitivity is commonly approximated
with the \acf{ECS}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation}
Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
from the external forcing $F$ and the climate feedback parameter $\lambda$. The
steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
simulation that is not in equilibrium by extrapolation with a linear regression
\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
  regression}, the global and annual mean net \ac{TOA} radiation $N$ versus the
change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
for energy leakage and remove any model drift present in the control climate, a
linear fit of the corresponding pre-industrial control run is subtracted from
the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
combination of \cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the \yintercept{} of this linear regression
($F_\mathtxt{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \mathtxt{ECS} = -\frac{F_\mathtxt{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
equilibrium climate sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

A further metric describing the climate sensitivity is the \ac{TCR}. In
contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
the Earth system but describes the transient response of an evolving climate.
Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a $1 \unit{\%}$
per year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}).
For this, the annual mean \acp{GSAT} are averaged over a 20 year period
centered at the time of the \ac{CO2} doubling (years \range{61}{80} when the
first year is indexed with $1$). To account for model drift, the annual mean
changes in \ac{GSAT} are calculated relative to a corresponding pre-industrial
control simulation smoothed with a linear fit that considers 140 model years
(length of the \onepctcotwo{} simulation). An illustration of that calculation
is shown in \cref{fig:02:tcr}.

@@

In the fast domain of the global carbon cycle, reservoir turnover times range
from seconds to millennia. In contrast to the slow domain, the carbon exchange
fluxes within the fast domain of the carbon cycle are much higher. One major
group of exchange processes in the fast domain connects the atmosphere and the
terrestrial biosphere. \ac{CO2} is removed from the atmosphere by plant
photosynthesis with about $120 \unit{\GtCyr{}}$ \autocite{Ciais2013}. This
process is also known as \ac{GPP}. The carbon fixed into plants can be released
back into the atmosphere by autotrophic (plant) and heterotrophic (soil
microbial and animal) respiration and additional disturbance processes like
fires \autocite{Ciais2013}. Since the land \ac{CO2} uptake by photosynthesis
occurs only during the growing season, whereas respiration occurs nearly all
year, the larger amount of vegetation in the Northern Hemisphere (due to the
larger land mass) gives rise to a seasonal cycle of the atmospheric \ac{CO2}
concentration \autocite{Keeling1995}. This seasonal cycle reflects the phase of
the global carbon cycle and shows a maximum of the atmospheric \ac{CO2}
concentration in the Northern Hemisphere winter (net \ac{CO2} flux into
atmosphere due to respiration) and a minimum during the Northern Hemisphere
summer (net \ac{CO2} flux into the land due to photosynthesis). Another major
carbon exchange process connects the atmosphere and the ocean. Atmospheric
\ac{CO2} is exchanged with the surface ocean through gas exchange, which is
driven by the partial \ac{CO2} pressure difference between the air and the sea
\autocite{Ciais2013}.

>>>>

An important metric describing climate sensitivity is the \emph{equilibrium
  climate sensitivity}. It is defined as the change in \ac{GSAT} after an
instantaneous doubling of the atmospheric \ac{CO2} concentration from
pre-industrial conditions once the climate system reaches radiative equilibrium
\autocite{Bindoff2013}. Being already used in one of the first assessments of
the anthropogenic climate change, the \emph{Charney Report} from 1979
\autocite{Charney1979}, the equilibrium climate sensitivity is one of the
oldest metrics describing climate change. However, in practice this traditional
definition of is not useful. Running fully-coupled \acp{ESM} into equilibrium
is computationally expensive as it would require thousands of model years
\autocite{Rugenstein2020}.

@@

For this reason, the equilibrium climate sensitivity is commonly approximated
with the \emph{\acf{ECS}}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation}
Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
from the external forcing $F$ and the climate feedback parameter $\lambda$. The
steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
simulation that is not in equilibrium by extrapolation with a linear regression
\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
  regression}, the global and annual mean net \ac{TOA} radiation $N$ versus the
change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
for energy leakage and remove any model drift present in the control climate, a
linear fit of the corresponding pre-industrial control run is subtracted from
the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
combination of \cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the \yintercept{} of this linear regression
($F_\mathtxt{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \mathtxt{ECS} = -\frac{F_\mathtxt{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
equilibrium climate sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

A further metric describing the climate sensitivity is the \emph{\ac{TCR}}. In
contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
the Earth system but describes the transient response of an evolving climate.
Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a $1 \unit{\%}$
per year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}).
For this, the annual mean \acp{GSAT} are averaged over a 20 year period
centered at the time of the \ac{CO2} doubling (years \range{61}{80} when the
first year is indexed with $1$). To account for model drift, the annual mean
changes in \ac{GSAT} are calculated relative to a corresponding pre-industrial
control simulation smoothed with a linear fit that considers 140 model years
(length of the \onepctcotwo{} simulation). An illustration of that calculation
is shown in \cref{fig:02:tcr}.

@@

In the fast domain of the global carbon cycle, reservoir turnover times range
from seconds to millennia. In contrast to the slow domain, the carbon exchange
fluxes within the fast domain of the carbon cycle are much higher. One major
group of exchange processes in the fast domain connects the atmosphere and the
terrestrial biosphere. \ac{CO2} is removed from the atmosphere by plant
photosynthesis with about $120 \unit{\GtCyr{}}$ \autocite{Ciais2013}. This
process is also known as \emph{\ac{GPP}}. The carbon fixed into plants can be
released back into the atmosphere by autotrophic (plant) and heterotrophic
(soil microbial and animal) respiration and additional disturbance processes
like fires \autocite{Ciais2013}. Since the land \ac{CO2} uptake by
photosynthesis occurs only during the growing season, whereas respiration
occurs nearly all year, the larger amount of vegetation in the Northern
Hemisphere (due to the larger land mass) gives rise to a seasonal cycle of the
atmospheric \ac{CO2} concentration \autocite{Keeling1995}. This seasonal cycle
reflects the phase of the global carbon cycle and shows a maximum of the
atmospheric \ac{CO2} concentration in the Northern Hemisphere winter (net
\ac{CO2} flux into atmosphere due to respiration) and a minimum during the
Northern Hemisphere summer (net \ac{CO2} flux into the land due to
photosynthesis). Another major carbon exchange process connects the atmosphere
and the ocean. Atmospheric \ac{CO2} is exchanged with the surface ocean through
gas exchange, which is driven by the partial \ac{CO2} pressure difference
between the air and the sea \autocite{Ciais2013}.

### 2021-03-01 ch07_conclusion/conclusion.tex
Finished chapter 01

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
ensemble. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations. In total we
assess eleven emergent constraints on \ac{ECS}, which are mostly related to
cloud feedbacks since these constitute the most important source of uncertainty
for \ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated
emergent constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5
ensemble, we can perform out-of-sample tests on the emergent constraints by
assessing whether they still hold for the \acs{CMIP}6 models. In this study,
which is shown in \cref{ch:05:paper_ecs} and already published in
\textcite{Schlund2020a}, we find a substantial reduction of skill for the
majority of emergent constraints when applied to the \acs{CMIP}6 ensemble in
comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
decrease of the coefficient of determination $R^2$ of the emergent relationship
and a decrease of the statistical significance using the null hypothesis that
there is no correlation between the predictor and \ac{ECS}. Moreover, the
corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
emergent constraints. Thus, \cref{enum:01:question_2}
(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
\enquote{no} for the \acs{CMIP}6 ensemble. The increased best estimates and
spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
\acs{CMIP}6. A possible reason for the reduced skill of the emergent
constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

@@

The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
ever increasing computational power and resources provided by modern
supercomputers has the potential for enormous improvements in climate modeling
and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
weighting approach, further possible applications of \ac{AI} and \ac{ML} in
climate science involve for example \ac{ML}-based parametrizations that are
learned from high-resolution climate models \autocite{Rasp2018}, the
\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
  Mansfield2020} or the learning of entire \acp{ESM} from observational
products \autocite{Geer2021}. Therefore, the foundation for innovative and
groundbreaking research in climate science has been laid, which will hopefully
help us in further understanding and mitigating one of the greatest challenges
for humankind today: climate change.

>>>>

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
ensemble. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. In total we assess eleven emergent constraints on
\ac{ECS}, which are mostly related to cloud feedbacks since these constitute
the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
  Flato2013}. Since all of the evaluated emergent constraints have been derived
on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, we can perform out-of-sample tests
on the emergent constraints by assessing whether they still hold for the
\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
already published in \textcite{Schlund2020a}, we find a substantial reduction
of skill for the majority of emergent constraints when applied to the
\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
skill is expressed as a decrease of the coefficient of determination $R^2$ of
the emergent relationship and a decrease of the statistical significance using
the null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
confidence) for almost all emergent constraints, resulting in values of
$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
be answered with a \enquote{no} for the \acs{CMIP}6 ensemble. The increased
best estimates and spreads resulting from the emergent constraints in
\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

@@

The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
ever increasing computational power and resources provided by modern
supercomputers has the potential for enormous improvements in climate modeling
and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
weighting approach, further possible applications of \ac{AI} and \ac{ML} in
climate science involve for example \ac{ML}-based parametrizations that are
learned from high-resolution climate models \autocite{Rasp2018}, the
\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
  Mansfield2020} or the learning of entire \acp{ESM} from observational
products \autocite{Geer2021}. Therefore, the foundation for innovative and
groundbreaking research in climate science for the near future has been laid,
which will hopefully help us in further understanding and fighting one of the
greatest challenges for humankind today: climate change.
